---
title: HighChart County Choropleth Map of County Poverty, 2015
author: John Hamilton Bradford
date: '2017-08-05'
slug: highchart-county-choropleth-map-of-county-poverty-2015
categories: []
tags: []
---

The following code downloads county-level poverty data in 2015 from the American Community Survey (ACS) via an API using the `acs` package.  I then create an interactive Highchart map using the package `highcharter`.

###Download ACS Poverty data via an API
```{r, tidy=TRUE, message=FALSE}

library(tibble)
library(dplyr)
library(tidyr)
library(fst)
library(highcharter)
library("viridisLite")
library(acs)

api.key.install(key="07e1aef2e1de18c0bd4ad888c8bd1295aa7400f9")
all_counties = acs::geo.make(state="*", county="*")
all_tracts = acs::geo.make(state=c(state.abb, "DC"), county="*", tract="*")


pov_total_2015 = acs::acs.fetch(endyear=2015, span=5, geography=all_counties,
                                table.number = "B17001", case.sensitive=F,
                                dataset="acs",
                                key = "07e1aef2e1de18c0bd4ad888c8bd1295aa7400f9")

df = tibble::as_tibble(pov_total_2015@estimate) %>% 
  dplyr::bind_cols(fips = paste0(as.character(pov_total_2015@geography$state), 
                                 as.character(pov_total_2015@geography$county))) %>% 
  dplyr::mutate(Poverty = 100*B17001_002/B17001_001) %>% 
  dplyr::select(fips, Poverty)


```


###Creating a HighChart Map
```{r, tidy=TRUE, message=FALSE}
pov <- df$Poverty
minVal=min(pov, na.rm=T)
maxVal=max(pov, na.rm=T)
val95 = quantile(pov, .95, na.rm=T)
val05 = quantile(pov, .05, na.rm=T)
valBins = 5
value_binned = round(maxVal/(valBins+1))

dclass <- data_frame(from = seq(minVal, maxVal-value_binned, by = value_binned),
                     to = seq(from[2], maxVal, by = value_binned),
                     color = substring(viridis(length(from), option = "C"), 0, 7))
dclass <- list_parse(dclass)


hc <- hcmap("https://code.highcharts.com/mapdata/countries/us/us-all-all.js", 
      data = df, value = "Poverty",
      joinBy = c("fips", "fips"), 
      name = "Poverty",
      dataLabels = list(enabled = FALSE, format = '{point.name}'),
      borderColor = "#FAFAFA", 
      borderWidth = 0.1,
      tooltip = list(headerFormat = "Poverty-",
                     valueDecimals = 2, 
                     valueSuffix = "%")) %>% 
  hc_title(text = "Percentage Living Below Poverty Level in 2015 (ACS)") %>% 
  hc_subtitle(text = "johnbradford.github.io") %>% 
  hc_mapNavigation(enabled = TRUE) %>% 
  hc_colorAxis(dataClasses = dclass) 
```


Once the chart is created, it has to be saved using the `htmlwidgets` package.  I saved it to a folder on my github page using the following code, and then use an iframe html tag to display it.  

```{r, eval=FALSE, tidy=TRUE}
library(htmlwidgets)
htmlwidgets::saveWidget(hc, file="us_poverty_map.html")
```


<iframe seamless src="/files/html_widgets/us_poverty_map.html" width="100%" height="700"></iframe>
